PhoneLets: offloading the phone off your phone for energy, cost and network load optimization

### 2014-06-17 paper.tex
changed a bit of discussion

\emph{PhoneLets} enable the functionality by exploiting remote SIM functionality of certain hardware, whereby a cellular module accepts to exchange commands with a SIM/UICC module over a separate communication link rather than direct, physical connection. \emph{PhoneLets} infrastructure establishes and manages secure communication channels between participating entities and allows a device to connect to a cellular network without a physical SIM module being present.

@@

\begin{figure}
\centering
\includegraphics[width=0.9\columnwidth]{figs/arch}
\vspace{-3mm}
\caption{Architecture}
\label{fig:arch}
\vspace{-5mm}
\end{figure}

@@

Our demo mobile client (Fig.~\ref{fig:client}) consists of a Telit UC864 cellular module connected to a raspberryPI as a controller, which also uses WiFi and BlueTooth dongles for local wireless connectivity. The controller runs a modified version of the Ofono open telephony stack\footnote{\url{https://ofono.org/}} and manages communication between the cellular module and multiple backends.

\begin{figure}[t!]
\centering
\includegraphics[width=0.9\columnwidth]{figs/sequence}
\vspace{-2mm}
\caption{Communication sequence between client, server and remote SIM card}
\label{fig:sequence}
\vspace{-5mm}
\end{figure}

@@

\begin{table}[t]
{
\small
\begin{tabular}{| l | c |}
\hline
  \textbf{Use case}         & \textbf{Avg. energy consumption (mA)}  \\ \hline
  Idle - airplane mode      & 5     \\ \hline
  Ifle - BlueTooth          & 6     \\ \hline
  Idle - WiFi               & 8     \\ \hline
  Idle - 3G                 & 10    \\ \hline
  Apps background - WiFi    & 27    \\ \hline
  BlueTooth SIM access      & 28    \\ \hline
  Apps background - 3G      & 40    \\ \hline
  WiFi tethering (idle)     & 80    \\ \hline
\end{tabular}
}
\vspace{-3mm}
\caption{Energy consumption of the different usage scenarios of a mobile device.}
\label{tab:energy}
\vspace{-3mm}
\end{table}

Energy consumption of a Samsung Galaxy S4 was measured\footnote{Using Monsoon Power Monitor \url{http://www.msoon.com/LabEquipment/PowerMonitor/}} in a number of scenarios. The different values are presented in Table.~\ref{tab:energy}. Energy consumption of an idle device with applications in background (\emph{Apps background}) considers a device with Google Apps, and Facebook installed and running in background and is consistent with our earlier observations~\cite{Aucinas:2013uk}. We have only run the test on a single WiFi network, therefore we expect the figure to vary, especially because our used network is a public one with some background traffic affecting energy consumption of the device.

@@

\begin{enumerate}
    \item Make a phone call to the mobile server (phone) and show that it is ringing
    \item Start sharing its SIM card with the client and show that the client is registered to the network with the same ID by making a phone call to it
    \item Start using SIM card in the SIM server and show that the client is registered to the network with a different ID by making a call to the other number.
\end{enumerate}

@@

As much as the telephone number is important, multiple providers have offered a service in the past whereby multiple SIM cards would be associated with the number and multiple handsets would ring on an incoming call, \eg Google Voice, however it does not solve the problem of an increasing number of cellular clients and billing depends on the exact agreement with the service provider. One cellular network in UK has even offered a service to make phonecalls from tablets and laptops over WiFi to its users using a special app\footnote{O2 TU Go \url{http://www.o2.co.uk/apps/tu-go}}.

@@

Finally, security is a major concern and the reason why secure elements are used by operators in the first place despite multiple proposals to move towards software-based SIMs. In our proposal, however, we still maintain the separate secure element and only replace the interface to it with an arguably secure alternative widely used in industry. Furthermore, past proposals of remote SIM usage have already included mobile payment solutions~\cite{subramanian2003sim}, which are even more risky than communication alone. Detailed security analysis is subject to future work, but we believe that any additional risk created by \emph{PhoneLets} is marginal.

@@

(1) \emph{Desk/table of approximately 1.5m, standard-width} to accommodate the demo. (2) \emph{Two standard UK power sockets}. (3) \emph{Two cellular network subscriptions (SIM cards)}, but are to be sources by the presenters, (4) WiFi AP for local communication. Setup time - up to an hour for connecting the demo up and verifying operation. Other equipment includes a laptop, two smartphones and the custom-build hardware described above.

>>>>

\emph{PhoneLets} enable the functionality by exploiting remote SIM functionality of existing hardware, whereby a cellular module accepts to exchange commands with a SIM/UICC module over a separate communication link rather than direct, physical connection. Some uses of such technology in the past have included automotive as well as electronic currency~\cite{subramanian2003sim}. \emph{PhoneLets} infrastructure establishes and manages secure communication channels between participating entities and allows a device to connect to a cellular network without a physical SIM module being present.

@@

\begin{figure}
\centering
\includegraphics[width=0.9\columnwidth]{figs/arch}
\vspace{-3mm}
\caption{Architecture}
\label{fig:arch}
\vspace{-2mm}
\end{figure}

@@

Our demo mobile client (Fig.~\ref{fig:client}) consists of a Telit UC864 cellular module connected to a raspberryPI as a controller, which also uses WiFi and BlueTooth dongles for local wireless connectivity. The controller runs a modified version of the Ofono open telephony stack and manages communication between the cellular module and multiple backends.

\begin{figure}[t!]
\centering
\includegraphics[width=0.9\columnwidth]{figs/sequence}
\vspace{-2mm}
\caption{Communication sequence between client, server and remote SIM card}
\label{fig:sequence}
\vspace{-4mm}
\end{figure}

@@

\begin{table}[t]
{
\small
\begin{tabular}{| l | c |}
\hline
  \textbf{Use case}         & \textbf{Energy consumption (mA)}  \\ \hline
  Idle - airplane mode      & 5     \\ \hline
  Ifle - BlueTooth          & 6     \\ \hline
  Idle - WiFi               & 8     \\ \hline
  Idle - 3G                 & 10    \\ \hline
  Apps background - WiFi    & 27    \\ \hline
  BlueTooth SIM access      & 28    \\ \hline
  Apps background - 3G      & 40    \\ \hline
  WiFi tethering (idle)     & 80    \\ \hline
\end{tabular}
}
\vspace{-2mm}
\caption{Energy consumption of the different usage scenarios of a mobile device.}
\label{tab:energy}
\vspace{-3mm}
\end{table}

Energy consumption of a Samsung Galaxy S4 was measured\footnote{Using Monsoon Power Monitor \url{http://www.msoon.com/LabEquipment/PowerMonitor/}} in a number of scenarios. The different values are presented in Table.~\ref{tab:energy}\footnote{Energy consumption provided as average current drawn because it is independent of current battery voltage}. Energy consumption of an idle device with applications in background (\emph{Apps background}) considers a device with Google Apps, and Facebook installed and running in background and is consistent with our earlier observations~\cite{Aucinas:2013uk}. We have only run the test on a single WiFi network, therefore we expect the figure to vary, especially because our used network is a public one with some background traffic affecting energy consumption of the device.

@@

\begin{enumerate}
    \item Make a phone call to the mobile server (phone) and show that it is ringing
    \item Start sharing its SIM card with the client and show that the client is registered to the network with the same ID by making a phone call to it
    \item Start using SIM card in the SIM server over WiFi and show that the client is registered to the network with a different ID by making a call to the other number.
\end{enumerate}

@@

\emph{PhoneLets} also offers management of part of an individual's ID - the phone number. In the past, multiple providers have offered a service in the past whereby multiple SIM cards would be associated with the number and multiple handsets would ring on an incoming call, \eg Google Voice, however it does not solve the problem of an increasing number of cellular clients and billing depends on the exact agreement with the service provider. One cellular network in UK has even offered a service to make phonecalls from tablets and laptops over WiFi to its users using a special app\footnote{O2 TU Go \url{http://www.o2.co.uk/apps/tu-go}}.

@@

Finally, questions that remain to be answered in future work include not only optimizations of the management and control protocols, but also the development of usage policies in multi-user environments and guaranteed isolation properties as well as adaptation of energy-aware routing protocols for distributed nodes.

@@

(1) \emph{Desk/table of approximately 2m, standard-width} to accommodate the demo. (2) \emph{Two standard UK power sockets}. (3) \emph{Two cellular network subscriptions (SIM cards)}, but are to be sources by the presenters, (4) WiFi AP for local communication. Setup time - up to an hour for connecting the demo up and verifying operation. Other equipment includes a laptop, two smartphones and the custom-build hardware described above.
